---
title: "Cornhole_2021_stats"
author: ""
date: "2023-03-04"
output:
html_document: default
---

```{r, warning = FALSE, message=FALSE}
library(readr)
library(dplyr)
library(scales)
library(DT)
library(kableExtra)
library(ggplot2)
```

# Data Cleaning

```{r}
# Pull Data

Master <- read.csv("master.csv", header=TRUE, skip=1)
Game_scores <- read.csv("Game_scores.csv", header=TRUE)
Round_scores <- read.csv("Round_scores.csv", header=TRUE)

Master$Match <- as.factor(Master$Match)
Master$Round <- as.factor(Master$Round)
Master$Throw_Number <- as.factor(Master$Throw_Number)
Master$Bag_Number <- as.factor(Master$Bag_Number)
```

```{r}
# Merge Data

Cornhole <- merge(Master, Round_scores, by.x = c("Match", "Round"), by.y = c("Game_Number", "Round_Number"))
Cornhole <- merge(Cornhole, Game_scores, by.x = "Match", by.y = "Game_Number")
Cornhole <- Cornhole%>%
  select(-c("Player_1.y", "Player_2.y"))
```


## Data Preview

```{r}
head(Master)
```


# Data Configuration


## Variable Breakdown by Throws

```{r}
Throw_breakdown <- Master%>%
  group_by(Throw_Number)%>%
  summarise(total = n(),
            in_hole_pct = round(sum(In_Hole)/total*100,1),
            on_board_pct = round(sum(On_Board)/total*100,1),
            missed_board_pct = round(sum(Missed_Board)/total*100,1),
            avg_Own_bags_dragged = sum(Own_Other_Bags_Dragged_In)/total,
            avg_Opponent_Bags_Dragged_In = sum(Opponent_Bags_Dragged_In)/total,
            avg_Own_Other_Bags_Knocked_Off = sum(Own_Other_Bags_Knocked_Off)/total,
            avg_Opponent_Other_Bags_Knocked_Off = sum(Opponent_Other_Bags_Knocked_Off)/total,
            Hole_Blocked_Before_Throw_pct = round(sum(as.numeric(Hole_Blocked_Before_Throw))/total*100,1),
            Hole_Blocked_After_pct = round(sum(Hole_Blocked_After_Throw)/total*100,1))

Throw_breakdown
```


## Basic Toss statistics by Thrower and Match

```{r, warning=FALSE, message=FALSE}
Throw_breakdown_by_game <- Master%>%
  group_by(Match, Thrower)%>%
  summarise(total = n(),
            in_hole_pct = round(sum(In_Hole)/total*100,1),
            on_board_pct = round(sum(On_Board)/total*100,1),
            missed_board_pct = round(sum(Missed_Board)/total*100,1),
            avg_Own_bags_dragged = sum(Own_Other_Bags_Dragged_In)/total,
            avg_Opponent_Bags_Dragged_In = sum(Opponent_Bags_Dragged_In)/total,
            avg_Own_Other_Bags_Knocked_Off = sum(Own_Other_Bags_Knocked_Off)/total,
            avg_Opponent_Other_Bags_Knocked_Off = sum(Opponent_Other_Bags_Knocked_Off)/total,
            Hole_Blocked_Before_Throw_pct = round(sum(as.numeric(Hole_Blocked_Before_Throw))/total*100,1),
            Hole_Blocked_After_pct = round(sum(Hole_Blocked_After_Throw)/total*100,1))

Throw_breakdown_by_game <- Throw_breakdown_by_game%>%
  select(1,2,4,5,6)

Throw_breakdown_by_game
```


## Visualizing a Round

```{r}
plot(Throw_breakdown$Throw_Number, Throw_breakdown$in_hole_pct, xlab = "Throw", ylab = "in hole percentage")
```

```{r}
plot(Throw_breakdown$Throw_Number, Throw_breakdown$Hole_Blocked_Before_Throw_pct, xlab = "Throw", ylab = "Hole Blocked For Throw")
```


# Analysis with Metrics


## Throw Skill

### Average Toss Score (ATS)

```{r, warning=FALSE, message=FALSE}
# By round average toss score
by_round_ATS <- Cornhole%>%
  group_by(Match, Round, Thrower)%>%
  summarise(ATS = (sum(On_Board) + 3*sum(In_Hole))/4)

# By game average toss score
by_game_ATS <- Cornhole%>%
  group_by(Match, Thrower)%>%
  summarise(total_throws = n(),
            ATS = (sum(On_Board) + 3*sum(In_Hole))/total_throws)%>%
  ungroup()
```


## Board Management

### Cluttering

#### Blocks Created (BC)

```{r, warning=FALSE, message=FALSE}
# By round blocks created
by_round_BC <- Cornhole%>%
  group_by(Match, Round, Thrower)%>%
  summarise(BC = sum(Strategic_Block_Created))
by_round_BC

# by game blocks created
by_game_BC <- Cornhole%>%
  group_by(Match, Thrower)%>%
  summarise(total_throws = n(),
            BC = sum(Strategic_Block_Created),
            NBC = BC/total_throws)
by_game_BC
```


#### Net Score after Block (NSAB)

```{r}
# Blocks in relation to winning

blocks_created_breakdown <- Cornhole%>%
  filter(Strategic_Block_Created == 1)%>%
  select(Match, Round, Throw_Number, Thrower, Round_Winner)%>%
  mutate(winning_block = ifelse(Round_Winner == Thrower, 1, 0))%>%
  group_by(Throw_Number)%>%
  summarise(total = n(),
            win_pct = sum(winning_block)/total)

blocks_table <- Cornhole%>%
  filter(Strategic_Block_Created == 1)%>%
  select(Match, Round, Throw_Number, Thrower, Round_Winner)%>%
  mutate(winning_block = ifelse(Round_Winner == Thrower, 1, 0))
```

```{r}
# Add final round scores to a table
final_round_scores <- Master%>%
  filter(Throw_Number == 8)%>%
  select(1,2,8,9)%>%
  rename("Player1_Final_Score" = Player1_Round_Score)%>%
  rename("Player2_Final_Score" = Player2_Round_Score)
```

```{r}
# Create a master table that includes player scores after a block (NSAB)
block_master <- Master%>%
  select(c(1:9,19:21))%>%
  mutate(Thrower_score_at_block = ifelse(Strategic_Block_Created == 1, ifelse(Thrower == Player_1, Player1_Round_Score, Player2_Round_Score), -1))%>%
  mutate(Opponent_score_at_block = ifelse(Strategic_Block_Created == 1, ifelse(Thrower == Player_1, Player2_Round_Score, Player1_Round_Score), -1))%>%
  merge(final_round_scores, by = c("Match", "Round"))%>%
  filter(Strategic_Block_Created == 1)%>%
  mutate(Thrower_Score_Diff = ifelse(Thrower == Player_1, Player1_Final_Score - Thrower_score_at_block, Player2_Final_Score - Thrower_score_at_block))%>%
  mutate(Opponent_Score_Diff = ifelse(Thrower == Player_1, Player2_Final_Score - Opponent_score_at_block, Player1_Final_Score - Opponent_score_at_block))%>%
  select(-c(3,8:12))%>%
  mutate(net_score_after_block = Thrower_Score_Diff - Opponent_Score_Diff)

mean(block_master$Thrower_score_at_block - block_master$Opponent_score_at_block)
mean(block_master$net_score_after_block)
```

```{r, warning=FALSE, message=FALSE}
## Block recovery score (NSAB)
by_game_NSAB <- block_master%>%
  group_by(Match, Thrower)%>%
  summarise(blocks = n(),
            total_NSAB = sum(net_score_after_block),
            NSAB_per_block = mean(net_score_after_block))

double_blocks <- blocks_table%>%
  group_by(Match, Round)%>%
  summarise(n = n())

by_round_NSAB <- block_master%>%
  group_by(Match, Round, Thrower)%>%
  summarise(blocks = n(),
            total_NSAB = sum(net_score_after_block),
            NSAB_per_block = mean(net_score_after_block))
```

```{r}
# Output number of blocks per throw and winning percentage if person created block on that round
blocks_created_breakdown
plot(blocks_created_breakdown$total, blocks_created_breakdown$win_pct)
text(blocks_created_breakdown$total, blocks_created_breakdown$win_pct, labels = blocks_created_breakdown$Throw_Number)
```


### Cleaning

#### Board Impact (BI)

```{r, warning=FALSE, message=FALSE}
# By round board impact
by_round_BI <- Cornhole%>%
  group_by(Match, Round, Thrower)%>%
  summarise(flat_BI = sum(Opponent_Other_Bags_Knocked_Off) + sum(Own_Other_Bags_Knocked_Off) + sum(Own_Other_Bags_Dragged_In) + sum(Opponent_Bags_Dragged_In),
            thrower_on_board = sum(On_Board))%>%
  group_by(Match, Round)%>%
  mutate(total_on_board = sum(thrower_on_board), 
         BI = flat_BI / total_on_board)%>%
  select(Match, Thrower, flat_BI, BI)
by_round_BI

# By game board impact
by_game_BI <- Cornhole%>%
  group_by(Match, Thrower)%>%
  summarise(flat_BI = sum(Opponent_Other_Bags_Knocked_Off) + sum(Own_Other_Bags_Knocked_Off) + sum(Own_Other_Bags_Dragged_In) + sum(Opponent_Bags_Dragged_In),
            thrower_on_board = sum(On_Board))%>%
  group_by(Match)%>%
  mutate(total_on_board = sum(thrower_on_board), 
         BI = flat_BI / total_on_board)%>%
  select(Match, Thrower, flat_BI, BI, total_on_board)
by_game_BI
```


#### Net Board Score (NBS)

```{r, warning=FALSE, message=FALSE}
# By round net board score
by_round_NBS <- Cornhole%>%
  group_by(Match, Round, Thrower)%>%
  summarise(flat_NBS = sum(Opponent_Other_Bags_Knocked_Off) - sum(Own_Other_Bags_Knocked_Off) + 2*sum(Own_Other_Bags_Dragged_In) - 2*sum(Opponent_Bags_Dragged_In))
by_round_NBS

# By game net board score
by_game_NBS <- Cornhole%>%
  group_by(Match, Thrower)%>%
  summarise(flat_NBS = sum(Opponent_Other_Bags_Knocked_Off) - sum(Own_Other_Bags_Knocked_Off) + 2*sum(Own_Other_Bags_Dragged_In) - 2*sum(Opponent_Bags_Dragged_In),
            thrower_on_board = sum(On_Board))%>%
  group_by(Match)%>%
  mutate(total_on_board = sum(thrower_on_board), 
         norm_NBS = flat_NBS / total_on_board)%>%
  select(Match, Thrower, flat_NBS, norm_NBS)
by_game_NBS
```


#### Overall Board Management Score (BMS)

```{r}
# Create overall board management score by combining NBS and NSAB metrics

# By match board management score
by_game_BMS <- by_game_BC%>%
  merge(by_game_BI, by = c("Match", "Thrower"))%>%
  merge(by_game_NBS, by = c("Match", "Thrower"))%>%
  merge(by_game_NSAB, by = c("Match", "Thrower"))%>%
  select(-BC)%>%
  mutate(won = as.factor(c(0,1,1,0,0,1,1,0,0,1,0,1,0,1,1,0,0,1)))%>%
  mutate(NBS_per_throw = flat_NBS / total_throws)%>%
  mutate(BMS = NBS_per_throw + NSAB_per_block)
by_game_BMS

# Board management score based on result (win/loss)
BMS_by_result <- by_game_BMS%>%
  group_by(won)%>%
  summarise(avg_BMS = mean(BMS),
            avg_NBS = mean(NBS_per_throw),
            avg_NSAB = mean(NSAB_per_block),
            var_NBS = sd(NBS_per_throw),
            var_NSAB = sd(NSAB_per_block),
            var_BMS = sd(BMS))
BMS_by_result
```


##### Prepare Table to plot BMS vs. ATS

```{r, warning=FALSE, message=FALSE}
# Manipulate table to only necessary columns, and add initials column
by_game_player_stats <- by_game_BMS%>%
  merge(by_game_ATS, by = c("Match", "Thrower", "total_throws"))
by_game_player_stats$Thrower_initials <- sapply(strsplit(by_game_player_stats$Thrower, " "), function(x) paste0(substr(x, 1, 1), collapse = ""))
by_game_player_stats$Thrower_initials <- ifelse(by_game_player_stats$won == 1,
                                                 paste0(by_game_player_stats$Thrower_initials, "*"),
                                                 by_game_player_stats$Thrower_initials)
```


# Summary Figures


## Tables

```{r}
by_game_stats <- by_game_ATS%>%
  merge(by_game_BC, by = c("Match", "Thrower"))%>%
  merge(by_game_BI, by = c("Match", "Thrower"))%>%
  merge(by_game_NBS, by = c("Match", "Thrower"))%>%
  merge(by_game_NSAB, by = c("Match", "Thrower"))%>%
  merge(Throw_breakdown_by_game, by = c("Match", "Thrower"))%>%
  mutate(won = as.factor(c(0,1,1,0,0,1,1,0,0,1,0,1,0,1,1,0,0,1)))%>%
  select(-total_throws.y)%>%
  rename("total_throws" = total_throws.x)%>%
  mutate(Toss_Score = ATS*total_throws)%>%
  mutate(NBS_per_throw = flat_NBS / total_throws)%>%
  mutate(BMS = NBS_per_throw + NSAB_per_block)%>%
  select(-BC)

by_game_stats
```


### Final Table (Pretty)

```{r}
# Cleaned up final output table
pretty_final_table <- by_game_stats%>%
  select(-c(norm_NBS, Toss_Score, total_on_board))%>%
  rename("Player" = Thrower, "Total Throws" = total_throws, "Average Toss Score" = ATS, "Blocks" = blocks, "Blocks per Throw" = NBC,  "Bags Affected" = flat_BI, "Board Impact" = BI, "Net Board Score" = flat_NBS, "NBS per Throw" = NBS_per_throw, "Net Score After Block" = total_NSAB, "NSAB per Block" = NSAB_per_block, "Board Management Score" = BMS, "Result" = won, "In Hole %" = in_hole_pct, "On Board %" = on_board_pct, "Missed Board %" = missed_board_pct)

pretty_final_table[, "Result"] = ifelse(pretty_final_table$Result == 1, "Winner", "Loser")

pretty_final_table1 <- pretty_final_table[, c(1,2,3,12,13,14,4,9,5,6,7,8,16,10,11,17,15)]

pretty_final_table1[, "Average Toss Score"] = round(pretty_final_table1$`Average Toss Score`, 3)
pretty_final_table1[, "Blocks per Throw"] = round(pretty_final_table1$`Blocks per Throw`, 3)
pretty_final_table1[, "NBS per Throw"] = round(pretty_final_table1$`NBS per Throw`, 3)
pretty_final_table1[, "NSAB per Block"] = round(pretty_final_table1$`NSAB per Block`, 3)
pretty_final_table1[, "Board Management Score"] = round(pretty_final_table1$`Board Management Score`, 3)

pretty_final_table1[, "Board Impact"] = percent(pretty_final_table1$`Board Impact`, .1)
pretty_final_table1[, "In Hole %"] = percent(pretty_final_table1$`In Hole %`/100, .1)
pretty_final_table1[, "On Board %"] = percent(pretty_final_table1$`On Board %`/100, .1)
pretty_final_table1[, "Missed Board %"] = percent(pretty_final_table1$`Missed Board %`/100, .1)

pretty_final_table1
```


#### Colorful pretty final table

```{r}
# create the table
pretty_final_table1 %>%
  kable("html", escape = F) %>%
  kable_styling(full_width = F, 
                bootstrap_options = "striped", 
                font_size = 16,
                stripe_color = "#F0F0F0") %>%
  add_header_above(c("Match Details" = 2, "Tossing Ability" = 5, "Board Management" = 9, "Result" = 1), 
                   background = "#9fc5e8") %>%
  # format header row
  row_spec(0, bold = T, 
           background = "#9fc5e8", 
           color = "black") %>%
  # format first column header
  column_spec(1:2, bold = T)
```


### Final Winner table

```{r}
# Table with the winner statistics for our created metrics
final_2way <- pretty_final_table %>%
  group_by(Result) %>%
  summarise(
    "Average Toss Score" = round(mean(`Average Toss Score`), 3),
    "In Hole %" = percent(mean(`In Hole %`)/100, .1),
    "On Board %" = percent(mean(`On Board %`)/100, .1),
    "Missed Board %" = percent(mean(`Missed Board %`)/100, .1),
    "Blocks per Throw" = round(mean(`Blocks per Throw`), 3),
    "NBS per Throw" = round(mean(`NBS per Throw`), 3),
    "Board Impact" = percent(mean(`Board Impact`), .1),
    "NSAB per Block" = round(mean(`NSAB per Block`), 3),
    "Board Management Score" = round(mean(`Board Management Score`), 3)
  )
final_2way
```

#### Colorful Final Winner table

```{r}
# create the table
final_2way %>%
  kable("html", escape = F) %>%
  kable_styling(full_width = F, 
                bootstrap_options = "striped", 
                font_size = 16,
                stripe_color = "#F0F0F0") %>%
  add_header_above(c("Result" = 1, "Tossing Ability" = 4, "Board Management" = 5), 
                   background = "#9fc5e8") %>%
  # format header row
  row_spec(0, bold = T, 
           background = "#9fc5e8", 
           color = "black") %>%
  # format first column header
  column_spec(1, bold = T)
```


### Final Game Scores (Pretty)

```{r}
# Cleaned up game scores table
Game_scores2 <- Game_scores%>%
  rename("Match" = Game_Number, "Player 1" = Player_1, "Player 2" = Player_2, "Player 1 Score" = Player_1_Final_Score, "Player 2 Score" = Player_2_Final_Score)
```

#### Colorful Final Game Scores table

```{r}
# create the table
Game_scores2 %>%
  kable("html", escape = F) %>%
  kable_styling(full_width = F, 
                bootstrap_options = "striped", 
                font_size = 16,
                stripe_color = "#F0F0F0") %>%
  # format header row
  row_spec(0, bold = T, 
           background = "#9fc5e8", 
           color = "black") %>%
  # format first column header
  column_spec(1, bold = T)
```


## Plots


### In Hole Freq Pretty Plot

```{r, warning=FALSE, message=FALSE}
# Define color palette
professional_colors <- c("#0072B2", "#f28e2b")

# Create ggplot object
in_hole_by_throw_plot <- ggplot(Throw_breakdown, aes(x = Throw_Number, y = in_hole_pct)) +
  
  # Add scatter plot with filled circles and alpha
  geom_point(size = 3, shape = 21, fill = professional_colors[1], alpha = 0.8, color = "white") +
  
  # Add a line plot with a thick line
  geom_line(size = 1.5, color = professional_colors[2]) +
  
  # Add x and y axis labels
  labs(x = "Throw", y = "In Hole Frequency") +

  # Change theme to use professional colors and add gridlines
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major = element_line(color = "gray", linetype = "dashed", size = 0.5),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f5f5f5", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "black", size = 1),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 18),
    plot.title = element_text(color = "black", size = 24, hjust = 0.5),
    legend.position = "none"
  )
 
# Add gridlines
in_hole_by_throw_plot +
  ggtitle("In Hole Frequency by Throw") +
  theme(
    panel.grid.major = element_line(color = "gray", linetype = "dashed", size = 0.5),
    panel.grid.minor = element_blank()
  )

#ggsave("InHole_Freq.png")
```


### Hole Blocked Pretty Plot

```{r, warning=FALSE, message=FALSE}
# Define color palette
professional_colors <- c("#0072B2", "#f28e2b")

# Create ggplot object
blocked_by_throw_plot <- ggplot(Throw_breakdown, aes(x = Throw_Number, y = Hole_Blocked_Before_Throw_pct)) +

  # Add scatter plot with filled circles and alpha
  geom_point(size = 3, shape = 21, fill = professional_colors[1], alpha = 0.8, color="white") +

  # Add a line plot with a thick line
  geom_line(size = 1.5, color = professional_colors[2]) +

  # Add x and y axis labels
  labs(x = "Throw", y = "Blocked Hole Frequency") +

  # Change theme to use professional colors and add gridlines
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major = element_line(color = "gray", linetype = "dashed", size = 0.5),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f5f5f5", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "black", size = 1),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 18),
    plot.title = element_text(color = "black", size = 24, hjust = 0.5),
    legend.position = "none"
  )

# Add gridlines
blocked_by_throw_plot +
  ggtitle("Blocked Hole Frequency by Throw") +
  theme(
    panel.grid.major = element_line(color = "gray", linetype = "dashed", size = 0.5),
    panel.grid.minor = element_blank()
  )

#ggsave("Block_Freq.png")
```

 
### BMS vs. ATS Plot

```{r}
# Create a ggplot object with the data and aesthetics
p <- ggplot(by_game_player_stats, aes(x = BMS, y = ATS, color = Match, label = Thrower_initials, fontface = ifelse(won == 1, "bold", "plain")))

# Add points to the plot
p <- p + geom_point(size = 3)

# Add x and y axis labels
p <- p + labs(x = "Board Management Score", y = "Average Throw Score")

# Set x-axis increments
p <- p + scale_x_continuous(breaks = seq(-2, 2, by = 0.5))

# Set y-axis increments
p <- p + scale_y_continuous(breaks = seq(1.4, 2.6, by = 0.2))

# Change theme to use professional colors and add gridlines
p <- p + theme_minimal(base_size = 16) +
  theme(
    panel.grid.major = element_line(color = "gray", linetype = "dashed", size = 0.5),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f5f5f5", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "black", size = 1),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 18),
    plot.title = element_text(color = "black", size = 24, hjust = 0.5),
    legend.position = "right",
    legend.title = element_text(face = "bold", size = 13)
  )

# Add labels to the points
p <- p + geom_text(nudge_x = 0.05, nudge_y = 0.05)

# Add a title to the plot
p <- p + ggtitle("BMS vs ATS by Match")

p

#ggsave("BMS_v_ATS.png")
```
